---
author: "Kharanshu Patel"
date: "10/6/2018"
---

``` 
library(ExPosition)
library(corrplot)
library(ggplot2)
library(InPosition)
library(dplyr)
library(PTCA4CATA)
library(data4PCCAR)
library(classInt)
library(Hmisc)
library(psych)
library(TInPosition)
library(RColorBrewer)

rm(list = ls())
graphics.off()
```
### Method: CA

Correspondence analysis (ca) is a generalized principal component analysis tailored for the analysis of qualitative data.The goal of correspondence analysis is to transform a data table into two sets of factor scores: One for the rows and one for the columns.

Rows and columns are displayed as points on the map whose coordinates are the factor scores and where the dimensions are called factors.

The factor scores of the rows and the columns have the same variance. Therefore, both rows and columns can be conveniently represented in one single map.

In correspondence analysis, a mass to each row and a weight to each column is assigned. The mass of each row reflects its importance in the sample. The mass of each row is the proportion of this row in the total of the table.
The weight of each column reflects its importance for discriminating between the variables. So the weight of a column reflects the information this columns provides to the identification of a given row.

### Dataset : Weekly Earnings by Race

The dataset contains information about the usual weekly earnings of people with various races and gender: White men, White women, Black men, Black women, Asian men, Asian women, Hispanic men, Hispanic women. These are the 8 rows. The 5 columns are 1st decile, 1st quartile, 2nd quartile, 3rd quartile and 9th decile. 

```{r data_set}

setwd('D://acn4/r studio class/')
WE <- read.csv('Weekly Earnings byRace(4).csv', row.names=1)

WE_data <- t(WE)
head(WE_data)
head(WE)

WE_dummy <- t(WE[(2:4), ])

WE_sup1 <- t(WE[1,]) 
WE_sup2 <- t(WE[6,]-WE[5,])

WE_col1 <- WE_dummy[,"1st quartile"]
WE_col2 <- WE_dummy[,"2nd quartile"] - WE_dummy[,"1st quartile"]
WE_col3 <- WE_dummy[,"3rd quartile"] - WE_dummy[,"2nd quartile"]
WE_col4 <- WE_data[,"Total people (in thousands)"] - WE_data[,"3rd quartile"]

WE_new <- cbind(WE_col1, WE_col2, WE_col3, WE_col4)
colnames(WE_new) = c("0-25%","25-50%","50-75%","75-100%")
colnames(WE_sup1) = c("0-10%")
colnames(WE_sup2) = c("90-100%")

WE_heat <- WE_new[ ,(1:3)]

```
## DESIGN

```{r design}

WE_DESIGN_gender <- rep(c("Men", "Women"),4)
WE_DESIGN_race <- rep(c("White", "Black", "Asian", "Hispanic"), each=2)

## COLOR

col4race2 = c("#00fcfe","#ff2ae3","#000969","#ab0205","#9800ff","#ff6205","#00688b","#ffe61c")

col4quart = c("greenyellow","green","green4","darkolivegreen")

```

### RESULTS

## COMPUTATIONS

# PCA
```{r PCA}

resPCA <- epPCA(DATA = WE_new,
                   scale = 'SS1', # Make to use 'SS1' rather than TRUE
                   graphs =  FALSE # TRUE first pass only
                    )
eigs <- resPCA$ExPosition.Data$eigs
PCA.plot.fi <- prettyPlot(resPCA$ExPosition.Data$fi, col = col4race2, display_names = TRUE)

PCAplot.fj <- prettyPlot(resPCA$ExPosition.Data$fj, col = col4quart, display_names = TRUE)
```
# Varimax 
```{r Varimax}

testVari    <- data4PCCAR::epVari(resPCA)

# Plot Varimax rotations

labels4Vari <- PTCA4CATA::createxyLabels.gen(1,2,
                                    lambda = testVari$rotated.eigs,
                                     tau = testVari$rotated.t)
baseMap.j.rot <- PTCA4CATA::createFactorMap(testVari$rotated.J,
                                  col.points   = col4quart,
                                   alpha.points =  .3,
                                   col.labels   = col4quart,
                                   title = 'Loadings Post Varimax')
# arrows
zeArrows.rot <- addArrows(testVari$rotated.J, color = col4quart)
# A graph for the J-set
c001.aggMap.J.rot <- baseMap.j.rot$zeMap_background + # background layer
                     baseMap.j.rot$zeMap_dots +
                     baseMap.j.rot$zeMap_text +  # dots & labels
                     zeArrows.rot + labels4Vari
print(c001.aggMap.J.rot)

# Plot the Rotated observations too
# a graph of the observations
baseMap.i.rot <- PTCA4CATA::createFactorMap(testVari$rotated.I,
  col.points = col4race2,
  col.labels = col4race2,
  display.labels = TRUE,
  alpha.points = .5,
  title = 'Factor Scores Post Varimax'
)
c002.Map.I.rot <- baseMap.i.rot$zeMap + labels4Vari

print(c002.Map.I.rot)
```
# Inference battery 
```{r Inference Battery}
resPCA.inf <- InPosition::epPCA.inference.battery(DATA = WE_new,
                   scale = 'SS1', # Make sure to use 'SS1' rather than TRUE
                   DESIGN = WE_DESIGN_race,
                   graphs =  FALSE # TRUE first pass only
                    )
```
## Group Analysis

# Bootstrap for CI:

```{r Bootstrap}
BootCube.Gr <- PTCA4CATA::Boot4Mean(resPCA$ExPosition.Data$fi, 
                                 design = WE_DESIGN_race,
                                 niter = 100,
                                 suppressProgressBar = TRUE)


# Bootstrap ratios 
bootRatios.Gr <- boot.ratio.test(BootCube.Gr$BootCube)

# eigenvalues: MonteCarlo Approach ----
 
random.eigen <- data4PCCAR::monteCarlo.eigen(X = WE_new, nIter = 100)

# eigenvalues: Bootstrap approach

bootstrap.eigen <- data4PCCAR::boot.eigen(WE_new, nIter = 100)
```

# CA analysis 

```{r CA analysis}

resCA.sym  <- epCA(WE_new, symmetric = TRUE)

# asymetric
resCA.asym <- epCA(WE_new, symmetric = FALSE)

# Supplementary Variables
WE_new_sup1 <- supplementaryCols(SUP.DATA = WE_sup1, res = resCA.sym)
WE_new_sup2 <- supplementaryCols(SUP.DATA = WE_sup2, res = resCA.sym)


# Factor Scores
Fj.a <- resCA.asym$ExPosition.Data$fj
Fi   <- resCA.sym$ExPosition.Data$fi
Fj   <- resCA.sym$ExPosition.Data$fj

```

# Constraints 

```{r Constraints}
# first get the constraints correct
constraints.sym <- minmaxHelper(mat1 = Fi, mat2  = Fj)
constraints.asym <- minmaxHelper(mat1 = Fi, mat2  = Fj.a)
constraints.sup <- minmaxHelper(mat1 = rbind(Fj, WE_new_sup1$fjj), 
                                mat2  = rbind(Fj, WE_new_sup2$fjj) )

```
## PLOTTING

# Heat Map(Not Correlation)

```{r Heat Map}
heatMapIJ.WE <- makeggHeatMap4CT(WE_new,
colorAttributes = col4quart,
colorProducts = col4race2,
fontSize.x = 15)
print(heatMapIJ.WE)

heatMapIJ.WE2 <- makeggHeatMap4CT(WE_heat,
colorAttributes = col4quart,
colorProducts = col4race2,
fontSize.x = 15)
print(heatMapIJ.WE2)

#dev.off()
```

## Scree Plot + Inference (Permutation Results)

```{r Scree Plot + Inference (Permutation Results)}

PlotScree(ev = eigs, 
       p.ev = resPCA.inf$Inference.Data$components$p.vals,
       title = 'Eigenvalues Inference',
       plotKaiser = TRUE
          )
```

# Base Map

```{r Base Map}
baseMap.i <- createFactorMap(Fi, constraints = constraints.sym,
                             col.points = col4race2,
                             col.labels = col4race2)
print(baseMap.i$zeMap)

baseMap.j <- createFactorMap(Fj, constraints = constraints.sym,
                             color.points = gplot::col2hex(col4quart),
                             color.labels = gplot::col2hex(col4quart),
                             pch = 11)
print(baseMap.j$zeMap)


```

## Factor Scores : Symmetric

```{r Factor Scores : Symmetric}
symMap  <- createFactorMapIJ(Fi,Fj,
                             col.points.i = col4race2,
                             col.labels.i = col4race2,
                             col.points.j = col4quart,
                             col.labels.j = col4quart,
                             pch.i = 15, pch.j = 19,
                             alpha.labels.i = 0.8,
                             alpha.labels.j = 0.8,
                             alpha.axes = 0.2,
                             col.axes = "#333434",
                             col.background = "#fff3e1"
                             )
labels4CA <- PTCA4CATA::createxyLabels.gen(1,2, lambda = resCA.sym$ExPosition.Data$eigs, tau = resCA.sym$ExPosition.Data$t)

map.IJ.sym <- symMap$baseMap + symMap$I_labels + symMap$I_points +
  symMap$J_labels + symMap$J_points + labels4CA +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

print(map.IJ.sym)


```

## Factor Scores : Asymmetric

```{r Factor Scores : Assymetric}
asymMap  <- createFactorMapIJ(Fi,Fj.a,
                              col.points.i = col4race2,
                              col.labels.i = col4race2,
                              col.points.j = col4quart,
                              col.labels.j = col4quart,
                              pch.i = 15, pch.j = 19,
                              alpha.labels.i = 0.8,
                              alpha.labels.j = 0.8,
                              alpha.axes = 0.2,
                              col.axes = "#333434",
                              col.background = "#f6e9e9"
                              )

map.IJ.asym <- asymMap$baseMap + asymMap$I_labels + 
  asymMap$I_points + asymMap$J_labels + 
  asymMap$J_points + labels4CA + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

print(map.IJ.asym)

```
## Plot Supplemenatary variables
```{r Supplementary Variables}

mapSup <- createFactorMapIJ(as.data.frame(WE_new_sup1$fjj), 
                            as.data.frame(WE_new_sup2$fjj),
                            col.points.i = "red1",
                            col.labels.i = 'red1',
                            col.points.j = 'red1',
                            col.labels.j = 'red1',
                            constraints = constraints.sup
)

# create an asymetric map with a supplementary row
map.I.sup.asym <- asymMap$baseMap + asymMap$I_labels + 
  asymMap$I_points +
  asymMap$J_labels + asymMap$J_points + 
  mapSup$I_labels + mapSup$I_points +
  labels4CA
print(map.I.sup.asym)

# Create a symmetric map with sup and correct constraints
map.IJ.sup.sym <- mapSup$baseMap + 
  symMap$I_labels + symMap$I_points +
  symMap$J_labels + symMap$J_points + 
  mapSup$I_labels + mapSup$I_points + 
  mapSup$J_labels + mapSup$J_points +
  ggtitle('Symmetric Map with Supplementary Elements') + 
  labels4CA
print(map.IJ.sup.sym)
```
## Contribution Bars

```{r Contribution Bars}
# compute the signed contributions
signed.ctrJ <- resPCA$ExPosition.Data$cj * sign(resPCA$ExPosition.Data$fj)
```

# Signed Contribution # 1

```{r Signed Contribution 1}
b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(col4quart), 
                         main = 'PCA on the WE_data Set: Variable Contribution 1 (Signed)',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b003.ctrJ.s.1)
```
# Signed Contribution 2

```{r Signed contribution 2}
b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                           threshold = 1 / NROW(signed.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(col4quart), # we need hex code
                           main = 'PCA on the WE_data Set: Variable Contribution 2 (Signed)',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b004.ctrJ.s.2)

```

## Unsigned Contributions

```{r Unsigned Contributions}
unsigned.ctrJ <- resPCA$ExPosition.Data$cj
```
# Unsigned Contribution 1
```{r Unsigned Contribution 1}
b003.ctrJ.u.1 <- PrettyBarPlot2(unsigned.ctrJ[,1],
                         threshold = 1 / NROW(unsigned.ctrJ),
                         font.size = 5,
                         color4bar = gplots::col2hex(col4quart), 
                         main = 'PCA on the WE_data set : Variable Contribution 1 (Unsigned)',
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b003.ctrJ.u.1)
```
# Unsigned Contribution 2
```{r Unsigned contribution 2}
 
b004.ctrJ.u.2 <- PrettyBarPlot2(unsigned.ctrJ[,2],
                           threshold = 1 / NROW(unsigned.ctrJ),
                           font.size = 5,
                           color4bar = gplots::col2hex(col4quart), # we need hex code
                           main = 'PCA on the WE_data Set: Variable Contribution 2(Unsigned)',
                           ylab = 'Contributions',
                           ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b004.ctrJ.u.2)

```
## Bootstrap Ratio Bars

```{r Bootstrap Ratios}
BR <- resPCA.inf$Inference.Data$fj.boots$tests$boot.ratios
```
# Dimension 1
```{r Dimension 1}
laDim = 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                        threshold = 2,
                        font.size = 5,
                   color4bar = gplots::col2hex(col4quart), # we need hex code
                   main = paste0(
                     'PCA on the WE_data Set: Bootstrap ratio ',laDim),
                  ylab = 'Bootstrap ratios'
                  #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
print(ba001.BR1)
```
# Dimension 2
```{r Dimension 2}
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            color4bar = gplots::col2hex(col4quart),
                            main = paste0(
                              'PCA on the WE_data Set: Bootstrap ratio ',laDim),
                ylab = 'Bootstrap ratios'
)
print(ba002.BR2)
```


### COnclusion


